---
title: 'Tidy Tuesday: Transit Costs Project'
author: ravic
date: '2021-01-04'
slug: tidy-tuesday-transit-costs-project
categories: []
tags:
  - tidy-tuesday
---

A #tidytuesday for transit costs.

<!--more-->

A #tidytuesday for transit costs.

```{r set-options, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set( message = FALSE, warning = FALSE )
options(scipen=999)
options(digits=4)
set.seed(722)
```
```{r load-packages, include=FALSE}
options(tidyverse.quiet = TRUE)
library("tidyverse")
require("scales", quietly=TRUE, warn.conflicts = FALSE)
require("glue", quietly=TRUE, warn.conflicts = FALSE)
require("patchwork", quietly=TRUE, warn.conflicts = FALSE)
suppressPackageStartupMessages(library("tidymodels"))
```
```{r plot-options, include=FALSE}
knitr::opts_chunk$set(dev.args=list(bg="transparent"))
theme_set(theme_light())
theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
             plot.background = element_rect(fill = "transparent", colour = NA))
```
```{r import-data, include=FALSE}
base_url <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/"
file_path <- "2021/2021-01-05/transit_cost.csv"
  
df_raw <- 
  readr::read_csv(
    paste0(base_url, file_path),
    col_types = cols(
      e = col_double(),
      country = col_character(),
      city = col_character(),
      line = col_character(),
      start_year = col_character(),
      end_year = col_character(),
      rr = col_double(),
      length = col_double(),
      tunnel_per = col_character(),
      tunnel = col_double(),
      stations = col_double(),
      source1 = col_character(),
      cost = col_double(),
      currency = col_character(),
      year = col_double(),
      ppp_rate = col_double(),
      real_cost = col_character(),
      cost_km_millions = col_double(),
      source2 = col_character(),
      reference = col_character()
    )
  )

readr::spec(df_raw)
```
```{r clean-data, include=FALSE}
require("httr", quietly=TRUE, warn.conflicts=FALSE)
require("countrycode", quietly=TRUE, warn.conflicts=FALSE)

df <- df_raw %>%
  rename(entry = e,
         tunnel_pct_complete = tunnel_per,
         country_code = country,
         currency_code = currency) %>%
  filter(!is.na(entry)) %>%
  mutate(tunnel_pct_complete = parse_number(tunnel_pct_complete) / 100,
         real_cost = na_if(parse_number(real_cost), 0)) %>%
  mutate(across(ends_with("_year"), ~ suppressWarnings(as.double(str_replace(.x, "\\D", ""))))) %>%
  mutate(across(ends_with("_year"), ~ if_else(.x < 1900, NA_real_, .x))) %>%
  mutate(across(starts_with("source"), factor)) %>%
  mutate(railroad_ind = as.logical(rr)) %>%
  mutate(tunnel_complete_ind = near(tunnel_pct_complete, 1)) %>%
  select(-rr, -tunnel_pct_complete)
```

```{r basic-features, include=FALSE}
df <- df %>%
  mutate(country_code = case_when(country_code == "UK" ~ "GB", 
                                  TRUE ~ country_code)) %>%
  mutate(country_name = map_chr(country_code, countrycode::countrycode, 
                                origin="iso2c", dest="country.name")) %>%
  mutate(country_name = factor(country_name)) %>%
  mutate(across(c(country_code, currency_code), factor)) 

df <- df %>%
  mutate(duration = end_year - start_year) 

df <- df %>%
  mutate(parsed_url = map(reference, parse_url)) %>% 
  mutate(hostname = map_chr(parsed_url, pluck, "hostname", .default=NA)) %>%
  select(-parsed_url) 

unfiltered <- df

df <- df %>%
  filter(tunnel_complete_ind,
         real_cost > 0,
         end_year < 2020) %>%
  mutate(log_real_cost = log1p(real_cost))
  
```
```{r split-data, include=FALSE}
splitter <- initial_split(df, prop = .8)
train <- training(splitter)
test <- testing(splitter)
```
```{r skim-data, eval=FALSE, include=FALSE}
library("skimr")
skimr::skim(df)
```
```{r correlations-pairs, include=FALSE, eval=FALSE}
library("ggcorrplot")
df %>% 
  select(where(is.numeric), -entry, -ends_with("year")) %>%
  filter(complete.cases(.)) %>%
  cor(.) %>%
  ggcorrplot(., hc.order = TRUE, legend.title=bquote(rho)) +
  theme_light() +
  labs(title="Correlation plot",
       y = NULL,
       x = NULL) +
  theme(axis.text.x = element_text(angle = 90, hjust=1))

```

```{r view-data, eval=FALSE, include=FALSE}
df %>%
  view(title="df")
```



Because of China's overwhelming investment in public transit, this dataset tells 
stories mostly of China. However, we're most interested in the United States 
story here. We'll caveat at the beginning that this is a pretty small sample,
of the transit records which are completed and have cost information as shown 
here (N = `r df %>% nrow()`):

```{r sample-size, echo= FALSE}
df %>% 
  mutate(country_name = fct_lump(country_name, n = 10)) %>%
  group_by(country_name) %>%
  summarise(n = n(),
            .groups="drop") %>%
  mutate(pct = n / sum(n)) %>%
  arrange(country_name == "Other", desc(n)) 
  
```


The distribution of the real costs across completed projects is roughly log-
normal. In judging each country's mean cost, we correspondingly used the
geometric mean, which shows the United States roughly equal to China and Japan.
However, the scatter indicates a more variance for the United States and Japan
with fewer, dispersed observations.

```{r country-plot, echo=FALSE, fig.width=8, fig.height=4}
real_cost_dist <- df %>%
  filter(tunnel_complete_ind, 
         real_cost > 0) %>%
  ggplot(aes(real_cost)) +
  geom_histogram(fill = "lightblue", bins=20) +
  scale_x_log10() +
  labs(title = "Distribution of real cost, log scale",
       x = NULL,
       y = NULL)

country_primed <- df %>%
  filter(tunnel_complete_ind,
         real_cost > 0) %>%
  mutate(country_name = fct_lump(country_name, n=10)) %>%
  group_by(country_name) %>%  
  summarise(geom_mean_cost = exp(mean(log(real_cost + 1))),
            .groups = "drop")

real_cost_scatter <- df %>%
  filter(tunnel_complete_ind,
         real_cost > 0) %>%
  mutate(country_name = fct_lump(country_name, n=10)) %>%
  left_join(country_primed, by = "country_name") %>%
  mutate(country_name = fct_reorder(country_name, geom_mean_cost)) %>%
  mutate(country_name = fct_relevel(country_name, "Other")) %>%
  ggplot(aes(real_cost, country_name)) +
  geom_point(alpha=.3) +
  geom_point(data = country_primed, aes(x=geom_mean_cost), color="blue", shape=4, size=2) +
  scale_x_log10() + 
  labs(title = "Real cost by country",
       subtitle = "Includes geometric mean real cost for each country",
       y = NULL,
       x = NULL) + 
  theme(
    panel.grid.minor = element_blank(), 
    panel.grid.major = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank())

real_cost_dist | real_cost_scatter
```

The next set of questions for those in the United States is, "Why are the costs
so high?" We could imagine that the projects in the United States are uniquely
complex (especially with such a small sample), or that they are overly long in
duration for their complexity -- or that they are expensive even taking time 
and project complexity into consideration.

Here, we fit a linear regression to all completed projects with real cost 
information, and see what we can determine.

```{r plot-stations-tunnels, eval=FALSE, include=FALSE}
df %>%
  filter(stations > 0,
         real_cost > 0,
         tunnel_complete_ind) %>%
  ggplot(aes(stations, real_cost)) +
  geom_point(alpha=.2) + 
  geom_smooth(formula=y ~ x, method="loess", se=TRUE) +
  scale_x_continuous(trans="log1p", breaks=seq(10, 60, 10)) +
  scale_y_continuous(trans="log1p", breaks=10 ^ seq(2, 5)) +
  coord_cartesian(xlim=c(1, 50)) +
  labs(title="Relationship between real cost and stations", 
       y = "Real cost",
       x = "Stations")

df %>%
  filter(tunnel > 0,
         real_cost > 0,
         tunnel_complete_ind) %>%
  ggplot(aes(tunnel, real_cost)) +
  geom_point(alpha=.2) + 
  geom_smooth(formula=y ~ x, method="loess", se=TRUE) +
  scale_x_continuous(trans="log1p", breaks=seq(10, 60, 10)) +
  scale_y_continuous(trans="log1p", breaks=10 ^ seq(2, 5)) +
  coord_cartesian(xlim=c(1, 50)) +
  labs(title="Relationship between real cost and tunnels", 
       y = "Real cost",
       x = "Tunnels")


```

```{r tidy-linear-model}
library("tidymodels")

linear_mod <- linear_reg() %>%
  set_engine("lm")

transit_recipe <- 
  recipe(log_real_cost ~ entry + length + duration + country_name + tunnel, 
         data = df) %>%
  update_role(entry, new_role="id") %>%
  step_filter(tunnel > 0, skip=FALSE) %>%
  step_naomit(duration, tunnel, country_name, length) %>%
  step_log(tunnel) %>%
  step_other(country_name, threshold = 10) %>%
  prep()

transit_workflow <- workflow() %>%
  add_recipe(transit_recipe) %>%
  add_model(linear_mod) 

baked_train <- transit_recipe %>%
  bake(new_data = df)

fitted_workflow <- transit_workflow %>%
  fit(data=df)

```


Again, the dataset we're looking at is very small. So the high $R^2$ value 
should be taken with a great deal of salt.

```{r inspect-tidy-linear-model, echo=FALSE}
require("broom", quietly=TRUE, warn.conflicts = FALSE)

options(scipen=0)
options(digits=3)

fitted_transit_linear <- fitted_workflow %>%
  pull_workflow_fit() %>%
  pluck("fit")
augmented <- broom::augment(fitted_transit_linear)
# broom::tidy(fitted_transit_linear)
broom::glance(fitted_transit_linear)
```

And here we examine the linear regression residuals to see if any observations
warrant further notice. And here, we do find a few that are interesting, even
with inclusion of length, tunnels (complexity proxy) and duration into our 
covariates.

```{r residuals-outliers, eval=TRUE, echo = FALSE, fig.width=8, fig.height=4}
suppressWarnings(require("ggfortify", quietly=TRUE, warn.conflicts = FALSE))

autoplot(fitted_transit_linear, alpha=.2, which=1:2) +
  theme_light() +
  theme(legend.position = "none") +
  theme(panel.background = element_rect(fill = "transparent", colour = NA),
        plot.background = element_rect(fill = "transparent", colour = NA))


augmented %>%
  bind_cols(baked_train %>% select(entry)) %>%
  left_join(df, by = "entry") %>%
  filter(.std.resid > 2) %>%
  mutate(line = str_sub(line, 1, 30)) %>%
  select(country_name=country_name.y, city, line, length = length.y, 
         tunnel = tunnel.y, duration = duration.x, real_cost)


```

<!-- Appendix: Unused code -->

```{r evaluate-linear-model, eval=FALSE, echo=FALSE}
library("tidymodels")

baked_test <- transit_recipe %>%
  bake(new_data = test)

predictions <-
  fitted_workflow %>%
  predict(new_data=test) %>%
  rename(prediction = .pred) %>%
  bind_cols(baked_test %>% select(log_real_cost)) %>%
  filter(is.finite(prediction)) %>%
  mutate(across(c(prediction, log_real_cost), exp))

eval_metrics <- metric_set(rmse, rsq)
eval_metrics(predictions, truth=log_real_cost, estimate=prediction)

```
```{r final-linear-model, eval=FALSE, echo=FALSE}
options(scipen=0)
final_fitted_lm <- transit_workflow %>%
  fit(data = df) %>%
  pull_workflow_fit() %>%
  pluck("fit") 
final_fitted_lm %>%
  tidy(conf.int = TRUE)
final_fitted_lm %>%
  autoplot(., alpha=.2, which=c(1, 2, 4, 6)) +
  theme_light() +
  theme(legend.position = "none") 
augmented <- broom::augment(final_fitted_lm)
baked_df <- transit_recipe %>%
  bake(new_data=df)


augmented %>%
  bind_cols(baked_df %>% select(entry)) %>%
  select(entry, starts_with(".")) %>%
  left_join(df %>% select(-length, -duration, -country_name, -tunnel, -length), by="entry") %>%
  filter(.std.resid > 2) 

baked_df %>%
  slice(c(39, 40, 90)) %>%
  select(entry) %>%
  left_join(df, by="entry")
```
```{r duration-model, eval=FALSE, echo=FALSE}
duration_model <- df %>%
  filter(end_year < 2020,
         complete.cases(.)) %>%
  mutate(country_name = fct_lump(country_name, n = 10)) %>%
  lm(duration ~ log1p(stations) + length + country_name, data=.) 

duration_model %>%
  broom::glance() 

duration_model %>%
  autoplot(., alpha=.4) +
  theme_light() 

df %>%
  filter(end_year < 2020,
         complete.cases(.)) %>%
  slice(c(35, 66, 30))
```

## Notes-to-self

[1]: https://yihui.org/knitr/options/
[2]: https://stackoverflow.com/questions/17152312/r-shiny-removing-ggplot2-background-to-make-it-transparent
[3]: https://gist.github.com/cboettig/5600558

1.  Making a transparent background for the plots was harder to track down than
    I would have thought. Yihui Xie's [detailed guide][1] has some description,
    but I needed to update the themes as well, as shown here[2] and here[3]














