---
title: Rolling with advantage
author: ravic
date: '2020-05-15'
slug: rolling-with-advantage
categories: []
tags:
  - riddler
---

From the Riddler, an extension to [Dungeons & Dragons](https://fivethirtyeight.com/features/can-you-find-the-best-dungeons-dragons-strategy/)

<!--more-->

```{r load_packages, include=FALSE}
library("tidyverse")
library("patchwork")
library("glue")
```

## Problem

From the Riddler, an extension to [Dungeons & Dragons](https://fivethirtyeight.com/features/can-you-find-the-best-dungeons-dragons-strategy/)

> The fifth edition of Dungeons & Dragons introduced a system of “advantage and disadvantage.” When you roll a die “with advantage,” you roll the die twice and keep the higher result. Rolling “with disadvantage” is similar, except you keep the lower result instead. The rules further specify that when a player rolls with both advantage and disadvantage, they cancel out, and the player rolls a single die. Yawn!

> There are two other, more mathematically interesting ways that advantage and disadvantage could be combined. First, you could have “advantage of disadvantage,” meaning you roll twice with disadvantage and then keep the higher result. Or, you could have “disadvantage of advantage,” meaning you roll twice with advantage and then keep the lower result. With a fair 20-sided die, which situation produces the highest expected roll: advantage of disadvantage, disadvantage of advantage or rolling a single die?

> Extra Credit: Instead of maximizing your expected roll, suppose you need to roll N or better with your 20-sided die. For each value of N, is it better to use advantage of disadvantage, disadvantage of advantage or rolling a single die?


## Answers
1.  Disadvantage-of-advantage gives the highest expected outcome: **11.17** vs. 
    10.5 for a straight roll, and 9.83 for advantage-of-disadvantage.
2.  If you need to roll N or better, you will want to use disadvantage-of-advantage
    for $1 \le N \le 13$, then resort to a straight roll for $14 \le N \le 20$.

## Analytic approach
Lots of order statistics. Order statistics on top of other order statistics.

### Formulation


Let's start with disadvantage of advantage. 

$$
\begin{aligned}
p_X(k) & = \frac{1}{20} \\
F_X(k) & = \frac{k}{20} \\
\\
A & = \max(X_1, X_2) = X_{max} \\
F_A(k) &= \left[F_X(k)\right]^2 \\
& = \frac{k^2}{20^2} \\
\\
D & = \min(A_1, A_2) = A_{min} \\
1 - F_D(k) &= \left[1 - F_A(k)\right]^2 \\
F_D(k) &= 1 - \left[1 - \frac{k^2}{20^2}\right]^2 \\
\end{aligned}
$$

Getting close. From here, we can get the probability mass function $p_D(k)$ for 
$D$, our random variable for disadvantage-of-advantage. And with that, we can
express the expectation $E[D]$ . 


$$
\begin{aligned}
p_D(k) & = F_D(k) - F_D(k-1) \\
E[D] & = \sum_k k \cdot p_D(k) \\
\end{aligned}
$$

```{r dis_of_adv, echo=FALSE}
faces <- 20
F_dis_of_adv <- function(k, faces) {
  1 - (1 - (k / faces)^2)^2
}
pmf_dis_of_adv <- function(k, faces) {
  F_dis_of_adv(k, faces) - F_dis_of_adv(k-1, faces) 
}
dis_of_adv_beats <- function(N, faces) {
  1 - F_dis_of_adv(N - 1, faces)
}
df <- tibble(k = seq_len(20)) %>%
  mutate(prob = pmf_dis_of_adv(k, faces),
         beats = dis_of_adv_beats(k, faces))
exp_adv_dis_fn <- function(faces) {
  sum(1:faces * pmf_dis_of_adv(1:faces, faces))
}
exp_dis_of_adv <- exp_adv_dis_fn(20)
```

### Results

Our expectation for disadvantage-of-advantage $E[D]$ works out to 
**`r {round(exp_dis_of_adv, 2)}`**. So, we get a little help above a 
straight roll with an expectation of 10.5.  That matches some intutition 
-- the advantage steps insulates us against the worst outcomes.  

We can plot the probability mass function, too. It's pretty hard to get 
either extreme -- the disadvantage step makes it hard to get 20, and the 
advantage makes it hard to get a 1. 

```{r plot_dis_of_adv, echo = FALSE}
df %>% 
  ggplot(aes(k, prob)) +
  geom_point(color="orange") +
  geom_segment(aes(xend=k, y=0, yend=prob), col="orange") +
  theme_light() +
  scale_y_continuous(labels = scales::percent_format(accuracy=1), 
                     breaks=seq(0, .08, .01), expand=c(0, 0)) +
  scale_x_continuous(breaks = seq(1, 20), expand = c(0.02, 0.02)) +
  labs(title = "Probability mass function of disadvantage-of-advantage", 
       x = "", y = "") +
  coord_cartesian(xlim=c(1, 20), ylim=c(0, .08)) +
  geom_hline(yintercept = .05, lty = 2, col="grey60") +
  theme(panel.grid= element_blank())
```


### What about advantage-of-disadvantage?

We could do the same order statistics for advantage-of-disadvantage. But 
that would be a drag. Instead, we'll just realize that it's the same 
probability mass function, just flipped over the line at $x = 10.5$. As 
[Missy Elliott would advise](https://genius.com/1227305), we need to "put 
our thing down, flip it and reverse it."  

Our expectation for advantage-of-disadvantage works out to $E[A^{\prime}] = \textbf{9.83}$.
And because the the advantage-of-disadvantage is consistently worse than a 
straight roll, we won't even bother considering it for the rest.

### What happens with more faces?

Looks a lot like $E[D] = E[X] + \frac{2}{3}$. That can't be a coincidence. 

We can leverage a second expression for expectation of a discrete distribution, 
and [Faulhaber's Formula](https://mathworld.wolfram.com/PowerSum.html) to get
an expression for $E[D]$. Let $m$ be the number of faces on the die, then:

$$
\begin{aligned}
E[D] & = \sum k \cdot p_D(k) \\
& = \sum_{k=0}^m [1 - F_D(k)] \\
& = \sum_{k=0}^m \left[1 - \frac{k^2}{m^2}\right]^2 \\
\\
\sum k^2 & = \frac{1}{3}m^3 + \frac{1}{2}m^2 + \frac{1}{6}m \\
\sum k^4 & = \frac{1}{5}m^5 + \frac{1}{2}m^4 + \frac{1}{3}m^3 - \frac{1}{30}m \\
\\
E[D] & = \frac{8}{15}m + \frac{1}{2} - \frac{1}{30m^3} \\
& \approx \frac{8}{15}m + \frac{1}{2}
\end{aligned}
$$

And now we can express the differences between the two expectations as a function
of the number of sides on the dice:
$$
\begin{aligned}
E[D] - E[X] & \approx [\frac{8}{15}m + \frac{1}{2}] - [\frac{m + 1}{2}] \\
& \approx \frac{1}{30}m \\
\end{aligned}
$$

And there's our $\frac{2}{3}$, hiding in plain sight.



---

### Extra credit

We already have most of the pieces in place for the extra credit. We can think
of the probability of getting $N$ or better as:
$$
\begin{aligned}
P(\text{outcome} \ge N) & = 1 - F(N - 1) \\
\\
F_X(k) & = \frac{k}{20} \\
F_D(k) &= 1 - \left[1 - \left(\frac{k}{20}\right)^2\right]^2 \\
\end{aligned}
$$

And then we can build a table, which matches our intuition from before -- that
disadvantage-of-advantage has a hard time reaching the extremes.

```{r extra_credit, echo = FALSE}
sdf <- tibble(N = seq_len(20)) %>%
  mutate(beats_straight = 1 - (N-1) / 20,
         beats_dis_of_adv = dis_of_adv_beats(N, faces)) %>%
  mutate(approach = ifelse(beats_dis_of_adv >= beats_straight, "dis-of-adv", "straight"))
sdf

```


---

### Second or third best?
What is disadvantage-of-advantage doing? We know it insulates us against the 
worst, and that it avoids the best too. Here's some simulation results -- mostly
second place, often third place, and even first place (with a tie) on occasion.

```{r simulation}
df <- tibble(roll = sample.int(n=20, size=10000, replace=TRUE)) %>%
  mutate(trial = floor((row_number() - 1)/ 4 + 1)) %>%
  mutate(pair = floor((row_number() - 1) / 2 + 1))

disadvantage_of_advantage <- df %>%
  group_by(trial, pair) %>%
  summarise(winner = max(roll)) %>%
  group_by(trial) %>%
  summarise(winner = min(winner)) 

df %>% 
  group_by(trial) %>%
  mutate(rank = rank(-1 * roll, ties.method="min")) %>%
  ungroup() %>%
  inner_join(disadvantage_of_advantage, 
             by = c(roll = "winner", trial = "trial")) %>%
  count(rank) %>%
  mutate(pct = n/sum(n))
  
```