---
title: Rolling with advantage
author: ravic
date: '2020-05-15'
slug: rolling-with-advantage
categories: []
tags:
  - riddler
---

From the Riddler, an extension to [Dungeons & Dragons](https://fivethirtyeight.com/features/can-you-find-the-best-dungeons-dragons-strategy/)

<!--more-->

```{r load_packages, include=FALSE}
library("tidyverse")
library("patchwork")
library("glue")
```

## Problem

From the Riddler, an extension to [Dungeons & Dragons](https://fivethirtyeight.com/features/can-you-find-the-best-dungeons-dragons-strategy/)

> The fifth edition of Dungeons & Dragons introduced a system of “advantage and disadvantage.” When you roll a die “with advantage,” you roll the die twice and keep the higher result. Rolling “with disadvantage” is similar, except you keep the lower result instead. The rules further specify that when a player rolls with both advantage and disadvantage, they cancel out, and the player rolls a single die. Yawn!

> There are two other, more mathematically interesting ways that advantage and disadvantage could be combined. First, you could have “advantage of disadvantage,” meaning you roll twice with disadvantage and then keep the higher result. Or, you could have “disadvantage of advantage,” meaning you roll twice with advantage and then keep the lower result. With a fair 20-sided die, which situation produces the highest expected roll: advantage of disadvantage, disadvantage of advantage or rolling a single die?

> Extra Credit: Instead of maximizing your expected roll, suppose you need to roll N or better with your 20-sided die. For each value of N, is it better to use advantage of disadvantage, disadvantage of advantage or rolling a single die?


## Analytic approach

**Lots of order statistics. Order statistics on top of other order statistics.**

Let's start with disadvantage of advantage. 

$$
\begin{aligned}
p_X(k) & = \frac{1}{20} \\
F_X(k) & = \frac{k}{20} \\
\\
A & = \max(X_1, X_2) = X_{max} \\
F_A(k) &= \left[F_X(k)\right]^2 \\
& = \left(\frac{k}{20}\right)^2 \\
\\
D & = \min(A_1, A_2) = A_{min} \\
1 - F_D(k) &= \left[1 - F_A(k)\right]^2 \\
F_D(k) &= 1 - \left[1 - \left(\frac{k}{20}\right)^2\right]^2 \\
\end{aligned}
$$

Getting close. From here, we can get the probability mass function (pmf) for 
$D$ for our random variable for disadvantage-of-advantage. And with that, we can
express the expectation $E[D]$ . 


$$
\begin{aligned}
p_D(k) & = F_D(k) - F_D(k-1) \\
E[D] & = \sum_k k \cdot p_D(k) \\
\end{aligned}
$$

```{r dis_of_adv, echo=FALSE}
faces <- 20
F_dis_of_adv <- function(k, faces) {
  1 - (1 - (k / faces)^2)^2
}
pmf_dis_of_adv <- function(k, faces) {
  F_dis_of_adv(k, faces) - F_dis_of_adv(k-1, faces) 
}
dis_of_adv_beats <- function(N, faces) {
  1 - F_dis_of_adv(N - 1, faces)
}
df <- tibble(k = seq_len(20)) %>%
  mutate(prob = pmf_dis_of_adv(k, faces),
         beats = dis_of_adv_beats(k, faces))
exp_dis_of_adv <- sum(seq_len(faces) * pmf_dis_of_adv(seq_len(faces), faces))
```

Some notes:

1. Our expectation for disadvantage-of-advantage $E[D]$ works out to 
**`r {round(exp_dis_of_adv, 2)}`**. So, we get a little help above a straight
roll with an expectation of 10.5.  That matches some intutition -- the advantage
steps insulates us against the worst outcomes.
2. Looks a lot like $E[D] = E[X] + \frac{2}{3}$. We'll need to come back to that.
3. We could do the same order statistics for advantage-of-disadvantage. But that's
a drag. Instead, we'll just realize that it's the same probability mass function,
just flipped over the line at $x = 10.5$.
As [Missy Elliott would advise](https://genius.com/1227305), we need to put our
thing down, flip it and reverse it.
4. We can plot the probability mass function, too. It's pretty hard to get either 
extreme -- the disadvantage step makes it hard to get 20, and the advantage 
makes it hard to get a 1. 

```{r plot_dis_of_adv, echo = FALSE}
df %>% 
  ggplot(aes(k, prob)) +
  geom_point(color="orange") +
  geom_segment(aes(xend=k, y=0, yend=prob), col="orange") +
  theme_light() +
  scale_y_continuous(labels = scales::percent_format(accuracy=1), 
                     breaks=seq(0, .08, .01), expand=c(0, 0)) +
  scale_x_continuous(breaks = seq(1, 20), expand = c(0.02, 0.02)) +
  labs(title = "Probability mass function of disadvantage-of-advantage", 
       x = "", y = "") +
  coord_cartesian(xlim=c(1, 20), ylim=c(0, .08)) +
  geom_hline(yintercept = .05, lty = 2, col="grey60") +
  theme(panel.grid= element_blank())
  # geom_point(y = .05, col = "lightblue", alpha=.4) +
  # geom_segment(aes(xend=k, y=0, yend=.05), col="lightblue", alpha=.4) +
  # geom_vline(xintercept = exp_dis_of_adv, lty=2, col="grey60") +
  # annotate("text", x = exp_dis_of_adv + .2, y = .085, 
  #          label=glue("Expected value:\n {round(exp_dis_of_adv, 2)}"),
  #          size=3, hjust=0, col="grey60") +


  
# df %>% 
#   ggplot(aes(k, beats)) +
#   geom_point(color="orange") +
#   theme_light() +
#   scale_y_continuous(labels = scales::percent_format(accuracy=1)) +
#   labs(x = "", y = "", title = "Probability of N or better")
```


We can get a bit closer to a closed-form solution for the expectation, using
[Faulhaber's formula](https://en.wikipedia.org/wiki/Faulhaber%27s_formula)

$$ 
\begin{aligned}
E[D] & = \sum_k k \cdot p_D(k) \\
& = \sum_k [1 - F_D(k)] \\
\\
& = \sum_k \left[1 - \left(\frac{k}{20}\right)^2\right]^2 \\
& = \sum_k \left[1 - 2\left(\frac{k^2}{20^2}\right) + \frac{k^4}{20^4}\right] \\
& = \sum_k 1 - \frac{2}{20^2}\sum_k k^2 + \frac{1}{20^4}\sum_k k^4 \\
\\
\sum_{k=1}^n k^2 & = \frac{n(n + 1)(2n + 1)}{6}\\
\sum_{k=1}^n k^4 & = \frac{n(n+1)(2n+1)(3n^2+3n-1)}{30} \\

\end{aligned}
$$

---

### Extra credit

```{r extra_credit, echo = FALSE}
sdf <- tibble(k = seq_len(20)) %>%
  mutate(beats_straight = 1 - (k-1) / 20,
         beats_dis_of_adv = dis_of_adv_beats(k, faces)) 
sdf

```

## Simulation

```{r simulation}
df <- tibble(roll = sample.int(n=20, size=10000, replace=TRUE)) %>%
  mutate(trial = floor((row_number() - 1)/ 4 + 1)) %>%
  mutate(pair = floor((row_number() - 1) / 2 + 1))

disadvantage_of_advantage <- df %>%
  group_by(trial, pair) %>%
  summarise(winner = max(roll)) %>%
  group_by(trial) %>%
  summarise(winner = min(winner)) 

adv_of_dis <- df %>%
  group_by(trial, pair) %>%
  summarise(winner = min(roll)) %>%
  group_by(trial) %>%
  summarise(winner = max(winner)) 

disadvantage_of_advantage %>%
  ggplot(aes(x = winner)) +
  geom_bar()

adv_of_dis %>%
  ggplot(aes(x = winner)) +
  geom_bar()

adv_of_dis %>%
  summarise(winner = mean(winner))

disadvantage_of_advantage %>%
  summarise(winner = mean(winner))

```

Now for advantage of disadvantage:

---


$$
\begin{aligned}
p_X(k) & = \frac{1}{20} \\
F_X(k) & = \frac{k}{20} \\
\\
D^{\prime} & = \min(X_1, X_2) = X_{min} \\
1 - F_D^{\prime}(k) &= \left[1 - F_X(k)\right]^2 \\
F_D^{\prime}(k) & = 1 - \left(1 - \frac{k}{20}\right)^2 \\
\\
A^{\prime} & = \min(D_1, D_2) = A_{max} \\
F_{A^{\prime}}(k) &= \left[F_{D^{\prime}}(k)\right]^2 \\
F_{A^{\prime}}(k) &=  \left[1 - \left(1 - \frac{k}{20}\right)^2\right]^2 \\
\end{aligned}
$$


And, again, we're close to the interesting stuff

$$
\begin{aligned}
p_{A^{\prime}}(k) & = F_{A^{\prime}}(k) - F_{A^{\prime}}(k-1) \\
P(k \ge N) & = 1 - F_{A^{\prime}}(N - 1) \\
\end{aligned}
$$

```{r adv_of_dis, echo=FALSE}
F_adv_of_dis <- function(k) {
  (1 - (1 - (k / 20))^2) ^2
}
pmf_adv_of_dis <- function(k) {
  F_adv_of_dis(k) - F_adv_of_dis(k-1) 
}
adv_of_dis_beats <- function(N) {
  1 - F_adv_of_dis(N - 1)
}
exp_adv_of_dis <- sum(seq_len(20) * pmf_adv_of_dis(seq_len(20)))
qdf <- tibble(k = seq_len(20)) %>%
  mutate(prob = pmf_adv_of_dis(k),
         beats = adv_of_dis_beats(k))
qdf %>% 
  ggplot(aes(k, prob)) +
  geom_point(color="orange") +
  theme_light() +
  scale_y_continuous(labels = scales::percent_format(accuracy=1)) +
  labs(x = "", y = "", title = "Probability of k") +
  geom_vline(xintercept = exp_adv_of_dis) 
exp_adv_of_dis
# |
# qdf %>% 
#  ggplot(aes(k, beats)) +
#  geom_point(color="orange") +
#  theme_light() +
#  scale_y_continuous(labels = scales::percent_format(accuracy=1)) +
#  labs(x = "", y = "", title = "Probability of N or better")
```

$$
\begin{aligned}
\sum_{k=1}^n k^2 & = \frac{n(n + 1)(2n + 1)}{6}\\
\sum_{k=1}^n k^4 & = \frac{n(n+1)(2n+1)(3n^2+3n-1)}{30} \\
\\
F_D(k) &= 1 - \left[1 - \left(\frac{k}{20}\right)^2\right]^2 \\
E[D] & = \sum 1 - F_D(k) \\
& = \sum \left(1 - \left(\frac{k}{n}\right)^2\right)^2 \\
& = \sum \left[1 - 2\left(\frac{k^2}{n^2}\right) + \frac{k^4}{n^4}\right] \\
& = \sum 1 - \frac{2}{n^2}\sum k^2 + \frac{1}{n^4} \sum k^4 \\
& = \frac{n(n+1)}{2} - \frac{2}{n^2}\left[\frac{n(n + 1)(2n + 1)}{6}\right] + \frac{1}{n^4}\left[\frac{n(n+1)(2n+1)(3n^2+3n-1)}{30}\right] \\
& = \frac{n(n+1)}{2} - \frac{2(n+1)}{3} + \frac{(n+1)}{5} \\
& = \frac{n + 1}{2} \cdot [(n - \frac{2(2n^2 + n)}{6n^2} + \dots] \\
& = \frac{n + 1}{2} \cdot [(n - \frac{2}{3} - \frac{1}{3n} + \dots] \\
\end{aligned}
$$

