---
title: Baseball asterisk
author: ravic
date: '2020-07-17'
slug: baseball-asterisk
categories: []
draft: true
tags:
  - riddler
---

<!--more-->



[1]:https://en.wikipedia.org/wiki/List_of_Major_League_Baseball_players_with_a_.400_batting_average_in_a_season
[2]:https://en.wikipedia.org/wiki/James%E2%80%93Stein_estimator



### Problem 

> From Taylor Firman comes an opportunity to make baseball history:  \ 
\
This year, Major League Baseball announced it will play a shortened 60-game
season, as opposed to the typical 162-game season. Baseball is a sport of
numbers and statistics, and so Taylor wondered about the impact of the season’s
length on some famous baseball records.  \
\
Some statistics are more achievable than others in a shortened season. Suppose
your true batting average is .350, meaning you have a 35 percent chance of
getting a hit with every at-bat. If you have four at-bats per game, what are
your chances of batting at least .400 over the course of the 60-game season? And
how does this compare to your chances of batting at least .400 over the course
of a 162-game season?


```{r load-packages, include = FALSE}
library("tidyverse")
library("scales")
library("janitor")
library("formattable")
library("Lahman")
```

### Answer

We can code up a quick look at the binomial distribution to see how likely we
would be to get a sample as extreme as the .400 record. 

```{r quick-p-value}
seasons <- c(60, 162)
at_bats_per_game <- 4

df <- tibble(seasons = c(60, 162)) %>%
  mutate(rarities = 1 - pbinom(.400 * seasons * at_bats_per_game, 
                               size = seasons * at_bats_per_game, 
                               prob = .350)) %>%
  mutate(rarities = formattable::percent(rarities, digits = 1))
df
```
Let's take a look at Ted Williams's record:

```{r ted-williams, echo = FALSE}
library("Lahman")

Batting %>%
  left_join(Master %>% select(playerID, nameFirst, nameLast), 
            by = "playerID") %>%
  mutate(batting_avg = round(H / AB, 3),
         ab_per_game = round(AB / G, 1)) %>%
  filter(nameLast == "Williams", nameFirst == "Ted") %>%
  select(nameFirst, nameLast, yearID, 
         H, AB, G, batting_avg, ab_per_game) %>%
  relocate(nameFirst, nameLast, H, AB, batting_avg) %>%
  arrange(yearID)
```

We can also take a look at those who achieved a .400 batting average in a single
season in the major leagues. Maybe most notable for our purposes are Ross Barnes
and Tuck Turner, who had fewer games, and bested .400 by quite a margin. (Ross
Barnes had previous .400 averages before the advent of the major leagues.)

The full list and comparison with their lifetime batting averages can be found
on [Wikipedia][1].

```{r list-of-400s, echo = FALSE}
Batting %>%
  left_join(Master %>% select(playerID, nameFirst, nameLast), 
            by = "playerID") %>%
  mutate(batting_avg = round(H / AB, 3),
         ab_per_game = round(AB / G, 1)) %>%
  filter(batting_avg >= .400) %>%
  filter(yearID >= 1876) %>%
  filter(G >= 60, AB >= 3 * 60) %>%
  arrange(yearID, desc(batting_avg)) %>%
  mutate(p_value = round(1 - pbinom(H, AB, prob = .350), 4)) %>%
  select(nameFirst, nameLast, yearID, 
         H, AB, G, batting_avg, ab_per_game, p_value) 
```




