---
title: 'Riddler: Square subdivisions'
author: ravic
date: '2021-01-08'
slug: riddler-square-subdivisions
categories: []
tags:
  - riddler
---
[1]: https://fivethirtyeight.com/features/can-you-cut-the-square-into-more-squares/

From the Riddler, a problem on [square subdivisions][1].

<!--more-->

```{r load-pacakges, include=FALSE, message=FALSE, warning=FALSE}
options(tidyverse.quiet = TRUE)
library("tidyverse")
```


## Problem

From the Riddler, a problem on [square subdivisions][1].

> From Rob Peacock comes a matter of squaring a multitude of squares: \
\
There are many ways to slice a big square into smaller squares (not necessarily
of equal size), so that the smaller squares don’t overlap, while still making up
the entire area of the big square. \
\
For example, you can slice the big square into four smaller squares, each a
quarter of the area of the big square. Or you could slice it into seven squares,
if you take one of those four squares and slice it into four yet smaller
squares. \
\
What whole numbers of squares can you not slice the big square into?


## Notes

One realization is that given a working number of squares, you can always
replace one of those squares with a set of four smaller squares. Another way to
articulate this is that for any valid $n$, $n + 3$ must also be valid.

The next realization is that you don't have to cut each square into quadrants,
but can instead tile the outside with small squares. If we take that approach,
we can find ourselves with 4, 6, or 8 squares based on our grid sizes:

```{r grids, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE,fig.height=3}
library("patchwork")

grid_plot <- function(s) {
  df <- expand.grid(x = 0:(s-1), y = 0:(s-1)) %>%
    mutate(width=1, height=1) %>%
    filter(x == 0 | y == (s-1)) %>%
    add_row(x = 1 + (s-2)/2, 
            y=  (s-2)/2, 
            width=(s-1), 
            height=(s-1)) %>%
    mutate(z = as.factor(if_else(x == 0 | y == (s-1), 1, 0))) 
  # default is compatible with geom_tile()
  ggplot(df, aes(x, y, width=width, height=height, fill = z)) + 
    geom_tile(color="black", alpha=.3) + 
    theme_minimal() +
    theme(legend.position = "none",
          panel.background = element_rect(fill = "transparent"),
          plot.background = element_rect(fill = "transparent", color = NA)) +
    scale_fill_manual(values = c("black", "darkgrey")) +
    scale_x_continuous(breaks=NULL, minor_breaks = NULL, labels=NULL) +
    scale_y_continuous(breaks=NULL, minor_breaks = NULL, labels=NULL) +
    labs(x= NULL,
         y = NULL) +
    coord_fixed()
}

grid_plot(1) | grid_plot(2) | grid_plot(3) | grid_plot(4)
```


If we combine those two ideas, we get four starters that we can build sequences
off of:

* 1
* 4: $\langle 4, 7, 10, 13, \dots \rangle$
* 6: $\langle 6, 9, 12, 15, \dots \rangle$
* 8: $\langle 8, 11, 14, 17, \dots \rangle$

## Answer

What's missing? The numbers **2, 3, and 5.** Turns out, there's no way to get
to those number of squares.





