---
title: 'Riddler: Anti-isoceles sets'
author: ravic
date: '2021-05-24'
slug: riddler-anti-isoceles-sets
categories: []
tags:
  - riddler
---


A Riddler challenge involving anti-isoceles sets.

<!--more-->
<script src="/rmarkdown-libs/twitter-widget/widgets.js"></script>

```{r load-packages, include=FALSE}
library("tidyverse")
```
```{r include=FALSE}
library("here")
best <- readr::read_csv(here("static/data/riddler", "isoceles-best.csv"))
iso_lengths <- readr::read_csv(here("static/data/riddler", "isoceles-set-lengths.csv"), col_types = cols())
```

### Problem

This feels like a NP-complete problem, where we are faced with too many possibilities to recursively exhaust for an exact answer for anything more than a small grid width $N$. Not quite sure yet which of the [Karp problems](https://en.wikipedia.org/wiki/Karp%27s_21_NP-complete_problems) most resembles this one, though. My approach here was simply to use the brute-force recursive approach, which allowed me to find solutions up to $N = 7$.  

While getting exact answers is possible for a limited range of grid-widths, the broader relationship seems hard to get a fix on.

### Limited solutions

The maximum size for the anti-isoceles set at each given grid-width is given here, with example plots below.  

```{r basic-summary, echo = FALSE}
best %>%
  group_by(grid_size) %>%
  summarise(anti_isoceles_set_size = max(size),
            .groups = "drop")
```


```{r example-solutions, echo = FALSE, fig.width = 6, fig.height = 6}
best %>%
  group_by(grid_size) %>%
  filter(idx == min(idx)) %>%
  ungroup() %>%
  ggplot(aes(x, y)) +
  geom_point() + 
  facet_wrap(~grid_size) +
  theme_light() +
  labs(subtitle = "Examples of largest anti-isoceles sets",
       x = NULL, y = NULL)
```



### Unique distances in the grid


```{r unique-distances, include=FALSE}
tbl <- tibble(
  grid_size = 1:7,
  non_isoceles_size = c(1, 2, 4, 6, 7, 9, 10),
  unq_distances = c(1, 3, 6, 10, 15, 20, 27))
```

Is the maximum size of the anti-isoceles set a function of the number of unique distances in the grid? The number of unique distances in the grid is a sequence with initial terms `r tbl$unq_distances`, and discussed as [OEIS A047800](https://oeis.org/A047800). 

The OEIS entry includes this conjecture from Vaclav Kotesovec (after Landau and Erdős), which shows a mostly quadratic term from the triangular sequence muted somewhat.

$a(n) \sim c \cdot \frac{n^2}{\sqrt{\log(n)}}, c = 0.79$


### Unique distances and relationship to largest anti-isoceles set

We might pose a possible relationship between the root of the number of unique distances and the size of the largest anti-isoceles set for a given grid width $N$. Basically, that the relationship is roughly linear with $N$, but takes into account that muting term above. Here, we plot these values, fitting a square root curve to the data.

```{r unique-distances-plot, fig.width=6, fig.height=6*5/8, echo = FALSE}
tbl %>%
  ggplot(aes(unq_distances, non_isoceles_size)) +
  geom_point() +
  geom_smooth(method="lm", formula=y ~ sqrt(x), se = F, color="black", 
              size=.5, lty=2) +
  labs(subtitle = paste0("Relationship between largest anti-isoceles set ",
                         "and unique distances in the grid"),
       x = "Unique distances",
       y = "Size of anti-isoceles set") +
  scale_y_continuous(breaks=seq(0, 10, 2), minor_breaks=0:10)  +
  theme_light() 
```


### Distribution of anti-isoceles set sizes

Looking beyond just the largest anti-isoceles set, it seems like that perhaps we can model the distribution at play here, using a Conway-Maxwell-Poisson (CMP) distribution. Does that suggest the a range of likely maximum values for each grid size?  

```{r distributions, fig.width = 6, fig.height = 6, echo = FALSE}
iso_lengths %>%
  ggplot(aes(size, sets)) +
  geom_col(fill = "lightblue") +
  theme_light() +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  facet_wrap(~grid_size, scales="free_y") +
  labs(subtitle = "Distribution of anti-isoceles set sizes",
       x = NULL, y = NULL) +
  scale_x_continuous(breaks=1:10, minor_breaks=NULL)
```

***

```{r related-circles, echo = FALSE, message = FALSE, warning = FALSE}
library("tweetrmd")
tweet_embed("https://twitter.com/9378282/status/1402665337418436614")
```


***

### Code

```{python, eval=FALSE}
import numpy as np
import pandas as pd
from scipy.spatial.distance import cdist


def unique_distances(grid_size: int):
    """Get number of unique distances in coordinate grid"""

    coords = [[x, y] for x in list(range(grid_size))
              for y in list(range(grid_size))]
    distances = cdist(coords, coords, metric="sqeuclidean")
    return len(np.unique(distances))


def validate(current):
    """Validate whether current set of points contains no isoceles triangles"""

    distances = cdist(current, current, metric="sqeuclidean")
    valid = np.all(
        np.apply_along_axis(
            lambda x: np.all(np.unique(x, return_counts=True)[1] <= 1),
            axis=1, arr=distances
        ))
    return valid


def construct(current, candidates):
    """Construct set of non-isoceles points"""

    if not candidates:
        return []

    with_new_item = current + [candidates[0]]
    if validate(with_new_item):
        return (
            construct(current, candidates[1:]) +
            construct(with_new_item, candidates[1:]) +
            [with_new_item]
        )
    return construct(current, candidates[1:])


def process_grid(grid_size: int):
    """Process grid with side length grid_size"""

    coords = [[x, y] for x in list(range(grid_size))
              for y in list(range(grid_size))]
    results = construct([], coords)
    return np.max([len(result) for result in results])


def main_control(max_size: int):
    """Loop through grids and process each"""

    max_sizes = [process_grid(idx) for idx in in list(range(max_size))]
    print(max_sizes)

if __name__ == "__main__":
    main_control(max_size=10)
```


<!--Appendix-->

```{r linear-relationship, include = FALSE, echo = FALSE, fig.width = 8, fig.width = 5}
best_summary <- best %>%
  group_by(grid_size) %>%
  summarise(anti_isoceles_set_size = max(size),
            .groups = "drop")

best_summary %>%
  ggplot(aes(grid_size, anti_isoceles_set_size)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, se = F, color = "black", 
              size = .5, lty = 2) +
  theme_light() +
  labs(subtitle = "Largest anti-isoceles set size vs. grid width",
       x = "Grid width",
       y = "Size of anti-isoceles set")
```


```{r linear-model, include = FALSE, fig.width = 8, fig.height = 5}
library("broom")
linear_model <- 
  best_summary %>%
  lm(data = ., formula = anti_isoceles_set_size ~ grid_size) 

summary(linear_model)
slope <- linear_model %>%
  broom::tidy() %>%
  filter(term != "(Intercept)") %>%
  pull(estimate)

library("broom")
new_tbl <- tibble(
  grid_size = 8:10,
  unq_distances = c(34, 42, 51)
)
model <- lm(data = tbl, formula = non_isoceles_size ~ sqrt(unq_distances))
summary(model)
model %>%
  broom::augment(newdata=new_tbl)
```




```{r include=FALSE, eval=FALSE}
best %>%
  group_by(grid_size) %>%
  summarise(num_unqs = n_unique(idx))
```




