---
title: 'Makeover Monday: Unpaid Labor'
author: ravic
date: '2020-04-06'
slug: makeover-monday-unpaid-labor
categories: []
tags:
  - makeover-monday
---

```{r load_packages, include=FALSE}
library("tidyverse")
library("janitor")
```


```{r import_data, include=FALSE}
df <- readr::read_csv("https://query.data.world/s/dgm572ybtukupkkld2ghdr4h2qvm6r") %>%
  janitor::clean_names()

```

```{r clean_data, include=FALSE}
df <- df %>% 
  mutate(gender = factor(tolower(gender))) %>%
  mutate_at(vars(time_use, age, survey_availability), factor) %>%
  mutate(start_year = as.numeric(str_sub(year, 0, 4)))

# Some countries provide data with regional breakdown.
# We will filter these countries out of the analysis.
countries_with_area <- df %>%
  filter(!is.na(area)) %>%
  pull(country) %>%
  unique()

df <- df %>%
  filter(!(country %in% countries_with_area)) %>%
  select(everything(), -area)

```

```{r skim_data, include=FALSE}
library("skimr")
skimr::skim(df)
```

### Spotlight: United States
The United States has steady data from 2003-2015 in this dataset (perhaps
from the Census Time Use Survey?). 

The timeframe includes some significant events (recoveries from 9/11, 2008 
financial shocks). The 2008 shock does seem visible in the men's paid labor
metrics -- without recovery -- but without any concurrent changes in men's
unpaid labor or women's labor metrics.

It's hard to gauge from the survey results, but it does seem like there is
some reduction in women's reported unpaid labor into early part of 2010's.
Is this simply choosing to spend less time here?

#### Ideas:
1.  "Pull your weight". A comparison of weight to unpaid labor by gender.
2.  "Bring home the bacon." 
3.  Rosie the Riveter is about women's paid labor. Timely brief (COVID vs. WW2)?

```{r usa_spotlight, echo = FALSE}
usa_detail <- df %>% 
  filter(country == "United States of America") %>% 
  arrange(start_year, gender) %>%
  select(start_year, gender, everything(), 
         -year, -survey_availability, -country, -age) %>% 
  mutate(time_use = fct_recode(
    time_use,
    unpaid = "Unpaid domestic, care and volunteer work",
    paid = "Paid and subsistence work"))

  
usa_gathered <- usa_detail %>% 
  pivot_wider(
    id_cols=start_year, 
    names_from=c(gender, time_use), 
    values_from=average_time_hours)

usa_detail %>%
  ggplot(aes(x=start_year, y=average_time_hours)) + 
  geom_line() +
  facet_grid(gender ~ time_use) + 
  theme_light() +
  labs(
    title = "Time use: Paid vs. unpaid, Male vs. Female",
    subtitle = "Country: USA",
    x = "",
    y = "Avg. time (hrs)"
  ) +
  scale_x_continuous(
    limits = c(2002, 2016),
    labels = scales::number_format(accuracy = 1, big.mark = ""),
    breaks = seq(2002, 2016, 2)) + 
  theme(axis.text.x = element_text(angle = 90))
```

