---
title: 'Tidy Tuesday: GDPR Violations'
author: ravic
date: '2020-04-21'
slug: tidy-tuesday-gdpr-violations
categories: []
tags:
  - tidy-tuesday
---


<!--more-->

### Background

[Overview from TidyTuesday](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-04-21/readme.md)

```{r load_packages, include=FALSE}
library("tidyverse")
library("janitor")
library("glue")
library("scales")
library("ggrepel")
library("formattable")
library("kableExtra")
```
```{r import_data, include=FALSE}
base_url <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/"
vdf_raw <- readr::read_tsv(
  glue(base_url, "2020/2020-04-21/gdpr_violations.tsv"),
  na = c("Not available.", "Not known", "0", "Not available", "Not disclosed", "Unknown")) %>%
  janitor::clean_names() 
tdf <- readr::read_tsv(glue(base_url, "2020/2020-04-21/gdpr_text.tsv")) %>%
  janitor::clean_names()
```
```{r clean_data, include=FALSE}
library("lubridate")
vdf <- vdf_raw %>%
  mutate_at(vars(name, authority, controller, type), factor) %>%
  mutate(date = na_if(mdy(date), ymd("1970-01-01")))
  
# Need to clean up `controller`. Different types (individual vs. company), 
# different variants on company name (e.g. Google vs. Google, Inc), and 
# different regions (Vodafone España vs. Vodafone Romania)
```
```{r data_overview, include=FALSE}
library("skimr")
skimr::skim(vdf)
vdf %>% view()
```


```{r key_events, fig.width=8, include=FALSE}
vdf %>%
  filter(!is.na(date), !is.na(price)) %>%
  ggplot(aes(x = date, y = price)) +
  geom_point(color = "lightblue") + 
  theme_light() +
  scale_y_log10(labels = scales::dollar_format(prefix = "\u20ac")) +
  labs(
    x = "",
    y = "",
    title = "Amount of GDPR fine (Eur.) over time"
  ) 
```
#### Examining biggest fines



Notes:

1.  Separate individuals and companies? Many small-value individual fines.
2.  Limited early enforcement. (The July 2018 fine in Portugal must have 
    really pissed somebody off.)
3.  Limited late-stage enforcement against individuals, small-value fines.
4.  Test early rapid enforcement deterrent via inter-arrival waiting times.

```{r reverse_date_axis, include=FALSE}
# https://stackoverflow.com/questions/43625341/reverse-datetime-posixct-data-axis-in-ggplot
c_trans <- function(a, b, breaks = b$breaks, format = b$format) {
  a <- scales::as.trans(a)
  b <- scales::as.trans(b)

  name <- paste(a$name, b$name, sep = "-")

  trans <- function(x) a$trans(b$trans(x))
  inv <- function(x) b$inverse(a$inverse(x))

  scales::trans_new(name, trans, inverse = inv, breaks = breaks, format=format)

}
rev_date <- c_trans("reverse", "date")
```

```{r timeline_plot, fig.height=8, fig.width=8, echo = FALSE}
vdf <- vdf %>%
  mutate(rank_fine = rank(-price)) 

vdf %>%
  filter(!is.na(date), !is.na(price)) %>%
  filter(price > 100000) %>%
  ggplot(aes(y = date, x = price)) +
  geom_jitter(color = "lightblue", show.legend = FALSE, height = 0, alpha = .7, width = .2) + 
  theme_light() +
  scale_x_continuous(
    breaks = seq(0, 5e7, 1e7),
    minor_breaks = seq(0, 5e7, 1e7),
    position = "top", 
    labels = scales::label_dollar(
      scale = 1e-6, accuracy = .1, 
      suffix = "",  prefix = "\u20ac")) +
  scale_y_continuous(
    trans = rev_date, labels = date_format("%b %Y"),
    breaks = ymd(c("2018-07-01", "2019-01-01", "2019-07-01", "2020-01-01", "2020-07-01"))) + 
  labs(
    x = "",
    y = "",
    title = "Amount of GDPR fine since enforcement began (2018-05-25)",
    subtitle = glue("Fines in millions of Euros (MM{euro})", euro="\u20ac"), 
    size = "Fine"
  ) +
  theme(axis.line.y.left = element_blank(),
        axis.line.y.right = element_blank(),
        panel.border = element_blank(),
        panel.grid.minor = element_blank()) +
  geom_text_repel(data = subset(vdf, price > 1e7), 
                  aes(label = glue("[{n}]", n=as.character(rank_fine))), hjust=0, 
                  segment.color = "grey", point.padding = 1,
                  size=rel(3)) 
  


```




```{r biggest_fines, echo = FALSE}
vdf %>% 
  arrange(desc(price)) %>%
  filter(price > 1e7) %>%
  arrange(desc(price)) %>%
  mutate(rank = rank(-price)) %>%
  select(rank, controller, date, fine=price, summary) %>%
  mutate(fine = scales::dollar(fine, scale = 1/1e6, accuracy=.1, prefix="\u20ac")) %>%  
  rename(`fine (MM €)` = fine) %>%
  knitr::kable() %>% 
  column_spec(column = 3, width = "100px") %>%
  column_spec(column = 4, width = "100px", )
```


---

### Further research

1.  The `summary` information is rich. Potentially understand the violation 
    without the text of the articles?
2.  Associations between violations? 
