---
title: Riddler Classic - GOAT Monty Hall
author: ravic
date: '2020-04-24'
slug: riddler-classic-goat-monty-hall
categories: []
tags:
  - riddler
---

From the Riddler, ["Now suppose Monty changes the rules."](https://fivethirtyeight.com/features/can-you-beat-the-goat-monty-hall-problem/)

<!--more-->

```{r load_packages, include = FALSE}
library("tidyverse")
library("glue")
library("formattable")
```

## Problem

From the Riddler, ["Now suppose Monty changes the rules."](https://fivethirtyeight.com/features/can-you-beat-the-goat-monty-hall-problem/)

> The Monty Hall problem is a classic case of conditional probability. In the original problem, there are three doors, two of which have goats behind them, while the third has a prize. You pick one of the doors, and then Monty (who knows in advance which door has the prize) will always open another door, revealing a goat behind it. It’s then up to you to choose whether to stay with your initial guess or to switch to the remaining door. Your best bet is to switch doors, in which case you will win the prize two-thirds of the time.

> Now suppose Monty changes the rules. First, he will randomly pick a number of goats to put behind the doors: zero, one, two or three, each with a 25 percent chance. After the number of goats is chosen, they are assigned to the doors at random, and each door has at most one goat. Any doors that don’t have a goat behind them have an identical prize behind them.

> At this point, you choose a door. If Monty is able to open another door, revealing a goat, he will do so. But if no other doors have goats behind them, he will tell you that is the case.

> It just so happens that when you play, Monty is able to open another door, revealing a goat behind it. Should you stay with your original selection or switch? And what are your chances of winning the prize?

## Answer
We're about to find that if Monty's opened a door, standing pat has a $\mathbf{\frac{3}{8}}$
probability of winning vs. changing doors, which has a $\mathbf{\frac{1}{2}}$ chance of 
winning. We're going with the **change doors** approach.


## Formulation

What's the probability that Monty can even open a door?

1.  Monty can't open a door to show a goat if there are no goats.
2.  Monty can't open a door if there's only one goat, and you chose that door.

$$
\begin{aligned}
P(D) & = \sum_{k=0}^3 p_G(k) \cdot P(D \mid G = k) \\
& = \frac{1}{4} \sum_{k=0}^3 P(D \mid G = k) \\
& = \frac{1}{4} \left[ 0 + \frac{2}{3} + 1 + 1 \right] \\
& = \frac{2}{3} \\
\end{aligned}
$$
The more interesting question is, "Given that Monty opened the door, what are 
the probabilities that there are 1, 2, or 3 goats?" (Just like the original 
Monty Hall problem, opening the door to show a goat provides a lot of 
information.)

$$
\begin{aligned}
P(G = k \mid D) & = \frac{P(D \mid G = k) \cdot P(G = k)}{P(D)} \\
& = \frac{1}{4} \cdot \frac{3}{2} \cdot P(D \mid G = k) \\
& = \frac{3}{8} \cdot P(D \mid G = k) \\
\\
P(G = 0 \mid D) & = 0 \\
P(G = 1 \mid D) & = \frac{3}{8} \cdot \frac{2}{3} = \frac{1}{4} \\
P(G = 2 \mid D) & = \frac{3}{8} \cdot 1  = \frac{3}{8} \\
P(G = 3 \mid D) & = \frac{3}{8} \cdot 1  = \frac{3}{8} \\
\end{aligned}
$$

Win probabilities for each of these scenarios is pretty easy to establish:

* If there were only one goat, and Monty showed it to you, then you've already won. 
* If there were three goats, you've already lost. :(
* And if there were two goats, we'd consult the original Monty Hall problem. :)  

Which we can lay out as:  

$$
\begin{aligned}
P(W \mid D, G = 0)  & = 0 \\
P(W \mid D, G = 1)  & = 1 \\
P(W \mid D, G = 2)  & = \begin{cases}
  \frac{1}{3} && \text{stand pat} \\
  \frac{2}{3} && \text{changing door} \\
  \end{cases} \\
P(W \mid D, G = 3)  & = 0 \\
\end{aligned}
$$



What's the probability that you will win **standing pat**, assuming Monty has opened a door?

$$
\begin{aligned}
P(W | D) & = \sum_{k=1}^3 P(G = k \mid D) \cdot P(W \mid D,\,G = k)  \\
& = \frac{1}{4} \cdot 1 + \frac{3}{8} \cdot \mathbf{\frac{1}{3}} + \frac{3}{8} \cdot 0 \\
& = \frac{1}{4} + \frac{1}{8}  \\
& = \mathbf{\frac{3}{8}} \\
\end{aligned}
$$

What's the probability that you will win **changing doors**, assuming Monty has opened a door?

$$
\begin{aligned}
P(W | D) & = \sum_{k=1}^3 P(G = k \mid D) \cdot P(W \mid D,\,G = k)  \\
& = \frac{1}{4} \cdot 1 + \frac{3}{8} \cdot \mathbf{\frac{2}{3}} + \frac{3}{8} \cdot 0 \\
& = \frac{1}{4} + \frac{1}{4}  \\
& = \mathbf{\frac{1}{2}} \\
\end{aligned}
$$


---

## Simulation

Simulation is a good idea with a Monty Hall problem.

```{r simulation}
set.seed(2020)
simulation_run <- function(stand_pat=TRUE) {
  # Place goats
  num_goats <- sample(0:3, 1)
  door_possibilities <- c(rep(1, num_goats), rep(0, 3 - num_goats))
  door_values <- sample(door_possibilities, 3, replace = FALSE)

    # Choose door
  idx_door_choice <- sample(1:3, 1)
  if (sum(door_values[-idx_door_choice]) == 0) {
    # Monty can't open another door!
    return(NA)
  }
  
  if (!stand_pat) {
    # One of the other doors had a goat that can be revealed
    # We will remove a goat-door, and our original choice.
    # Then we will revise our choice!
    
    doors = 1:3
    potential_goat_doors <- doors[doors != idx_door_choice & door_values[doors] == 1]
    goat_door <- potential_goat_doors[sample.int(length(potential_goat_doors), 1)]
    idx_door_choice <- doors[doors  != idx_door_choice & doors != goat_door]
  }
  
  # Did we win?
  is_winner <- (door_values[idx_door_choice] == 0)
  return(is_winner)
}

df <- crossing(
    trials = seq(1:100000),
    stand_pat = c(TRUE, FALSE)) %>%
  mutate(outcome = purrr::map_lgl(stand_pat, simulation_run))

df %>%
  group_by(stand_pat) %>%
  summarise(
    trials = n(), 
    door_open = sum(!is.na(outcome)),
    winning = sum(outcome, na.rm = TRUE)) %>%
  mutate(pct_winning = winning / door_open)
```
