---
title: 'Riddler: Ambiguous brackets'
author: ravic_
date: '2020-02-07'
slug: riddler-ambiguous-brackets
categories: []
fig_caption: yes
tags:
  - riddler
---
On the ambiguity in absolute values, from the 
[Riddler](https://fivethirtyeight.com/features/how-many-more-palindrome-dates-will-you-see/):

<!--more-->

On the ambiguity in absolute values, from the 
[Riddler](https://fivethirtyeight.com/features/how-many-more-palindrome-dates-will-you-see/):

> The pairings of other mathematical symbols can be more ambiguous. Take the absolute value symbols in Jim’s example, which are vertical bars, regardless of whether they mark the opening or closing of the absolute value. As Jim points out, |−1|−2|−3| has two possible interpretations: \
\
1. The two left bars are a pair and the two right bars are a pair. In this case, we have 1−2·3 = 1−6 = −5. \
\
2. The two outer bars are a pair and the two inner bars are a pair. In this case, we have |−1·2−3| = |−2−3| = |−5| = 5.\
\
How many different values can the expression |−1|−2|−3|−4|−5|−6|−7|−8|−9| have?

---


```{r load-packages, echo = FALSE}
options(tidyverse.quiet = TRUE)
library("tidyverse")
```

```{r define-functions, echo = FALSE}
library("tidyverse")
evaluate_options <- function(remnant, num_brackets, eqtn = "", num_used = 0, 
                             num_open = 0, pathway = "") {
  
  # Cannot leave the final expression unclosed
  if ((num_open > 0) &
      (num_used >= num_brackets)) {
    return(NULL)
  }
  
  # Cannot have a negative number of open brackets
  if (num_open < 0) {
    return(NULL)
  }
  
  # Terminal state
  # Evaluate equation string using:
  # https://stackoverflow.com/questions/1743698/evaluate-expression-given-as-a-string
  if ((num_open == 0) & 
      (num_used >= num_brackets)) {
    rslt_tibble <- tibble(equation = eqtn, pathway = pathway) %>%
      mutate(evaluation = eval(parse(text = eqtn)))
    return(rslt_tibble)
  }
  
  # Pull off first number and bracket 
  # Results include both opening and closing the brackets
  numeric <- str_extract(remnant, "[^|]+")
  new_remnant <- sub("[^|]+\\|", "", remnant)

  return(bind_rows(
    evaluate_options(
      # When opening a bracket, the new equation will be a concatenation:
      # existing equation + numeric token + "* abs("
      remnant = new_remnant, 
      num_brackets = num_brackets,
      eqtn = paste0(eqtn, numeric, "* abs("), 
      num_used = num_used + 1, 
      num_open = num_open + 1, 
      pathway = paste0(pathway, "O")),
    evaluate_options(
      # When closing a bracket, the new equation will be a concatenation:
      # existing equation + numeric token + ")"
      remnant = new_remnant, 
      num_brackets = num_brackets,
      eqtn = paste0(eqtn, numeric, ")"), 
      num_used = num_used + 1, 
      num_open = num_open - 1, 
      pathway = paste0(pathway, "C")))
    )
}
```

```{r eval-test-data, echo = FALSE}
# Adding 1 in front just for convenience
TEST_STRING <- "1|-1|-2|-3|-4|-5|-6|-7|-8|-9|"

answers <- evaluate_options(remnant = TEST_STRING, num_brackets = 10)
```


```{r summarise-results, echo = FALSE}
answer_summary <- answers %>% arrange(evaluation)
num_pathways <- nrow(answer_summary)
unq_answers <- unique(answers %>% pull(evaluation))
num_unq_answers <- length(unq_answers)

duplicate_answers <- unique(
  answer_summary %>% 
    add_count(evaluation) %>% 
    filter(n > 1) %>% 
    pull(evaluation))

```


## Summary
There are `r {num_pathways}` possible expressions, and a prettier way to get
at that fact than this code-heavy approach. But unfortunately, different 
formulations can -- and do -- lead to the same answer. So we have no choice
but to evaluate all the paths. :(

In the end, there are only **`r {num_unq_answers}` unique answers**. 
(The duplicate answers are: `r {duplicate_answers}`.) The full list of paths, 
equations and answers follows at the end.


---

## Code

Upon encountering a bracket while reading the string, we have a choice: we 
can either open a new bracket or close an already open bracket. This code 
relies on a recursive solution to explore all those possible choices and
evaluate the resulting final expressions. 

We're using a nice trick in R, which allows us to easily evaluate a string
as an expression.


```{r ref.label = "define_functions", eval = FALSE, echo = TRUE}
```

---

## Results

```{r ref.label = "eval_test_data", eval = FALSE, echo = TRUE}
```

```{r display_raw, echo = FALSE}
answer_summary %>% 
  print(n = nrow(answer_summary))
```

---

## Appendix  


```{r pathway-diagram, echo = FALSE, fig.cap = "Pathway diagram", dev="svg"}
library(ggrepel)
df <- crossing(num_used = seq(0, 10), num_open = seq(0, 5))
qdf <- tibble(
  num_used = c(0, 1, 2, 2, 3, 3, 4, 4, 4, 5, 5, 5, 6, 6, 6, 7, 7, 8, 8, 9, 10),
  num_open = c(0, 1, 2, 0, 3, 1, 4, 2, 0, 5, 3, 1, 4, 2, 0, 3, 1, 2, 0, 1, 0),
  paths = c(42, 42, 28, 14, 14, 14, 5, 9, 5, 1, 4, 5, 1, 3, 2, 1, 2, 1, 1, 1, 1)
)  

df %>% 
  ggplot(aes(num_used, num_open)) + 
  scale_y_reverse(breaks = seq(0, 5), minor_breaks = NULL) + 
  scale_x_continuous(position = "top", 
                     breaks = seq(0, 10), minor_breaks = NULL) + 
  geom_segment(x = 0, y = 0, xend = 5, yend = -5, color = "grey80", alpha = .2, lty = 2) +
  geom_segment(x = 2, y = 0, xend = 6, yend = -4, color = "grey80", alpha = .2, lty = 2) +
  geom_segment(x = 4, y = 0, xend = 7, yend = -3, color = "grey80", alpha = .2, lty = 2) +
  geom_segment(x = 6, y = 0, xend = 8, yend = -2, color = "grey80", alpha = .2, lty = 2) +
  geom_segment(x = 8, y = 0, xend = 9, yend = -1, color = "grey80", alpha = .2, lty = 2) +
  geom_segment(x = 1, y = -1, xend = 2, yend = 0, color = "grey80", alpha = .2, lty = 2) +
  geom_segment(x = 2, y = -2, xend = 4, yend = 0, color = "grey80", alpha = .2, lty = 2) +
  geom_segment(x = 3, y = -3, xend = 6, yend = 0, color = "grey80", alpha = .2, lty = 2) +
  geom_segment(x = 4, y = -4, xend = 8, yend = 0, color = "grey80", alpha = .2, lty = 2) +
  geom_segment(x = 5, y = -5, xend = 10, yend = 0, color = "grey80", alpha = .2, lty = 2) +
  geom_point(color = "grey80", shape = '+', size = 3) + 
  labs(x = "Brackets used", y = "Open\nbrackets",
       title = "Number of pathways for given state (used, open brackets)") + 
  theme_light() +
  theme(panel.grid = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_blank(),
        axis.title.x = element_text(hjust = 0, vjust = 1),
        axis.title.y = element_text(angle = 0, vjust = 1, hjust = 1)) +
  geom_point(data = qdf) +
  geom_text_repel(data = qdf, aes(label = paths), nudge_x = .1, nudge_y = .1, size = 3) +
  coord_fixed() 
```



```{r generalized-string, echo = FALSE, eval = FALSE}
# A generalized approach to handle longer and longer strings
gen_bracketed_string <- function(num_pairs) {
  innards <- paste0(seq(-1, num_pairs * (-2) + 1, -1), collapse = "|")
  return(paste0("1|", innards, "|"))
}

df <- tibble(num_bracket_pairs = 1:10) %>%
  mutate(bracketed_string = purrr::map(num_bracket_pairs, gen_bracketed_string)) %>%
  mutate(answers = map2(bracketed_string, num_bracket_pairs * 2, evaluate_options)) %>%
  mutate(num_pathways = map_int(answers, nrow)) %>%
  mutate(num_unique_answers = map_int(answers, 
                                  function(x) { length(unique(x %>% pull(evaluation))) })) %>%
  mutate(pct_unique = num_unique_answers / num_pathways)

df %>%
  select(num_bracket_pairs, num_pathways, num_unique_answers, 
        pct_unique)
```



