---
title: Riddler Classic - Coin Flip to Freedom
author: ravic
date: '2020-05-04'
slug: riddler-classic-coin-flip-to-freedom
categories: []
tags:
  - riddler
---


<!--more-->

## Problem
From the [Riddler](https://fivethirtyeight.com/features/can-you-flip-your-way-to-freedom/),
can you flip your way to freedom?

> From Bart Wright comes a rhetorical question from a famed soliloquy, “To flip, or not to flip?”:

> You are locked in the dungeon of a faraway castle with three fellow prisoners (i.e., there are four prisoners in total), each in a separate cell with no means of communication. But it just so happens that all of you are logicians (of course).

> To entertain themselves, the guards have decided to give you all a single chance for immediate release. Each prisoner will be given a fair coin, which can either be fairly flipped one time or returned to the guards without being flipped. If all flipped coins come up heads, you will all be set free! But if any of the flipped coins comes up tails, or if no one chooses to flip a coin, you will all be doomed to spend the rest of your lives in the castle’s dungeon.

> The only tools you and your fellow prisoners have to aid you are random number generators, which will give each prisoner a random number, uniformly and independently chosen between zero and one.

> What are your chances of being released?

> Extra credit: Instead of four prisoners, suppose there are N prisoners. Now what are your chances of being released?


## Formulation
We're going to assume each prisoner is quietly analyzing the cutoff value $c$
that would maximize all their chances of escape. Intuitively, we'd expect a 
somewhat low value for $c$ -- escape feels more likely if fewer prisoners are
flipping coins.

$$
\begin{aligned}
P_F & = \sum_{k=1}^4 \binom{4}{k} c^k(1-c)^{4-k} \cdot \left(\frac{1}{2}\right)^k \\
& = \sum_{k=1}^4 \binom{4}{k} \left(\frac{c}{2}\right)^k(1-c)^{4-k}  \\
& = \sum_{k=\mathbf{0}}^4 \binom{4}{k} \left(\frac{c}{2}\right)^k(1-c)^{4-k} - (1 -c)^4 \\
& = \left(\frac{c}{2} + (1-c)\right)^4 - (1 - c)^4 \\
& = \left(1 - \frac{c}{2}\right)^4 - (1 -c)^4 \\
\end{aligned}
$$

Since we're using the binomial expansion, we can generalize this to any number
of prisoners, $n$.

$$
\begin{aligned}
P_F & = \left(1-\frac{c}{2}\right)^n - (1-c)^n
\end{aligned}
$$

What is the optimal value of the cutoff $c$? Let's take the derivative of $P_F$
with respect to $c$, and see what cutoff will maximize our chances.

$$ 
\begin{aligned}
P_F & = \left(1-\frac{c}{2}\right)^n - (1-c)^n \\
\frac{dP_F}{dc} & =  n \left(1-\frac{c}{2}\right)^{n-1}\left(-\frac{1}{2}\right) - n(1-c)^{n-1}(-1) \\
0 & = \left(\frac{1}{2}\right)^n \cdot (2-c)^{n-1} - (1-c)^{n-1} \\
(1 - c)^{n-1} &= \left(\frac{1}{2}\right)^n \cdot (2-c)^{n-1} \\
(1 - c)^{n-1} &= \left(\frac{1}{2}\right)^{n-1} \cdot \left(\frac{1}{2}\right) \cdot (2-c)^{n-1} \\
1 - c & = \frac{1}{2} \cdot \frac{1}{2}^{1/(n-1)} \cdot (2 - c) \\
\end{aligned}
$$

Let's define a variable $r$ for that unwieldy root, and use that to express the
cutoff that we're looking for:

$$
\begin{aligned}
r & = 2^{1/(1-n)} \\
2 - 2c & = 2r - cr \\
2 - 2r & = 2c - cr \\
c & = \frac{2 - 2r}{2- r} \\
\end{aligned}
$$

## Plotting results
As we increase the number of prisoners, the cutoff value $c$ that maximizes
our probability of freedom predictably drives down quickly.

What isn't quite as obvious is that the probability would converge to a value
close to **25%**.  Maybe we could do a binomial expansion and see something
geometric within it?

$$
\begin{aligned}
P_F & = \left(1-\frac{c}{2}\right)^n - (1-c)^n
\end{aligned}
$$


```{r load_packages, include=FALSE}
library("tidyverse")
```

```{r gen_solutions}
# library("tidyverse")
unwieldy_root <- function(n) { 2 ^ (1 / (1 - n)) }
best_cutoff <- function(n) { 
  r = unwieldy_root(n)
  (2 - 2 * r) / (2 - r)
}
win_prob <- function(n) {
  c <- best_cutoff(n)
  (1 - c/2)^n - (1 - c)^n
}

df <- tibble(n = seq(2, 10000)) %>%
  mutate(
    c = best_cutoff(n),
    p = win_prob(n)) 
```

```{r plot_solutions}
df %>%
  ggplot(aes(x = n, y = c)) +
  geom_point(color="lightblue") +
  theme_light() +
  labs(title = "Optimal cutoff for varying number of prisoners",
       x = "",
       y = "") +
  scale_x_continuous(breaks = seq(0, 100, 10), 
                     minor_breaks = NULL) +
  coord_cartesian(xlim = c(0, 100))
```

```{r}
df %>% 
  ggplot(aes(x = n, y = p)) +
  geom_point(color="lightblue") +
  theme_light() +
  labs(title = "Probability of freedom for varying number of prisoners",
       x = "",
       y = "") +
  scale_x_continuous(breaks = seq(0, 100, 10), 
                     minor_breaks = NULL) +
  scale_y_continuous(labels = scales::percent_format(accuracy=1),
                     breaks = seq(0.25, 0.33, .01),
                     minor_breaks = NULL) +
  coord_cartesian(xlim = c(0, 100)) +
  geom_hline(lty=2, yintercept = 0.25)
```

```{r log_relationship}
df %>%
  ggplot(aes(x = n, y = c)) +
  geom_point(color="lightblue") +
  scale_x_log10() +
  scale_y_log10() +
  theme_light() +
  labs(title = "Power law relationship of best cutoff with number of prisoners",
       x = "",
       y = "")

```


## Limit

https://math.stackexchange.com/questions/882741/limit-of-1-x-nn-when-n-tends-to-infinity
https://socratic.org/questions/what-is-the-derivative-of-2-x
https://www.wyzant.com/resources/lessons/math/calculus/limits/lhopital

