---
title: Extra point vs. two-point conversion
author: Ravi Chandrasekaran
date: '2020-01-20'
slug: extra-point-vs-two-point-conversion
categories: []
math: true
tags:
  - riddler
---


From the Riddler, a [proposal](https://fivethirtyeight.com/features/can-you-track-the-delirious-ducks/) 
for a new-and-improved post-touchdown extra-point system. 

## The defense
#### Simple strategies
One simple, uniformed strategy the defense could employ would be to simply
defend against the 2-point conversion all the time. What would the offense's
expected score be?

$$
E[S_{CONV}] = p_{XP} \cdot 1.0 + (1 - p_{XP}) \cdot 0.4 \\
$$


The other simple, uninformed strategy would be to always defend against the
extra point all the time. The expected score there would be:

$$
E[S_{XP}] = p_{XP} \cdot 0.9 + (1 - p_{XP}) \cdot .6 \\
$$

We can graph both simple strategies and the offense's expected score as a 
function of their selected probability of going for the extra point $p_{XP}$:

```{r load_pacakges, echo = FALSE}
options(tidyverse.quiet = TRUE)
library("tidyverse")
```


```{r expected_score, echo = FALSE}
# Expected score
exp_score <- function(prob_xp, defend_xp) {
  xp <- prob_xp * 1.0  * ifelse(defend_xp, .9, 1.0) 
  conv <- (1 - prob_xp) * 2.0 * ifelse(defend_xp, .6, .4)
  return(xp + conv)
}

# Simulation
df <- tibble(prob_xp = seq(0, 1, .01)) %>%
  mutate(
    score_defend_xp = exp_score(prob_xp, TRUE),
    score_defend_conv = exp_score(prob_xp, FALSE))
```


```{r visualise_results, echo = FALSE}
df %>%
  gather(key = "defense", value = "score", -prob_xp) %>%
  ggplot(aes(prob_xp, score, col=defense)) +
  geom_line() +
  theme_light() +
  theme(legend.position = "bottom") +
  labs(
    title = "Expected scores with different defenses",
    x = "Probability of extra point attempt",
    y = "Expected point per play"
  ) +
  scale_x_continuous(labels = scales::percent_format()) +
  geom_vline(xintercept = .8, lty = 3)

```


#### General strategy
Since the defense actually knows the offense's selected probability of
going for the extra point, they can select the strategy that minimizes 
the offense's total scoring. They can defend against the two-point 
conversion until they know that that the offense is very likely to go
for the extra point. 

Wha's the threshold? We can set the two scoring equations equal, and 
that will give us the threshold at which we should change tack: $p_{XP} = 0.8$.

$$
p_{XP} \cdot 1.0 + (1 - p_{XP}) \cdot 0.4 = p_{XP} \cdot 0.9 + (1 - p_{XP}) \cdot .6 \\
p_{XP} = 0.8
$$



**So, the defense should defend against the two-point conversion until it 
knows the offense is going to go for the extra point with at least a 80% 
probability.**

---

## The offense

#### General strategy
The offense knows that the defense knows the offense's strategy. (Ugh, that's 
a mouthful.) But using the same graph above, the maximum expected score under
the defense's approach will also be at $p_{XP} = 0.8$. Substituting into either
equation gives us an expected score of $E[S] = `r {.8 * 1.0 + (1 - .8) * 0.4}`$.

**So, the offense should consistently send in a probability of 0.8, expecting to
score 0.88 points per attempt.**

