---
title: 'Week ending 2021-07-04'
author: ravic
date: '2021-07-04'
slug: week-ending-2021-07-04
categories: []
tags:
  - week-in-review
params:
  end_date: "2021-07-04"
---

```{r load-packages, include = FALSE}
options(tidyverse.quiet = TRUE)
library("tidyverse")
require("tweetrmd", quietly = TRUE, warn.conflicts = FALSE)
```
```{r date-range, include= FALSE}
library("lubridate")
end_date = as.Date(params$end_date)
start_date = end_date - 6
```
```{r get-twitter-favorites, include = FALSE, cache = TRUE}
require("rtweet", quietly = TRUE, warn.conflicts = FALSE)

rtweet::get_token()
faves <- rtweet::get_favorites(user = "ravic_")
```
```{r list-tweets, include = FALSE, eval=FALSE}
# More todos on handling the text: 
# 1. Deal with internal quotes, smart quotes
# 2. Remove media URL's, substitute link URL's
# 3. Link to screen names, highlight text-internal screen names, maybe real names
# 4. De-emoji, ftfy
# 5. Shorten to single sentence?

require("tweetrmd", quietly = TRUE, warn.conflicts = FALSE)
require("glue", quietly = TRUE, warn.conflicts = FALSE)

recent_faves <- faves %>%
  select(user_id, status_id, created_at, screen_name, text) %>%
  filter(between(as.Date(created_at), start_date, end_date)) %>%
  mutate(url = tweetrmd::tweet_url(user_id, status_id),
         create_date = as.character(created_at, "%b %d"),
         text = str_replace_all(text, "[\r\n]" , ""),
         screen_name = str_to_lower(screen_name),
         text = str_replace_all(text, ".http", ". http")) 
recent_faves %>% 
  arrange(created_at) %>%  
  mutate(idx = row_number(),
         # command = glue('tweet_embed("{url}")'),
         formatted = glue::glue('# {idx}. "{text}," --{screen_name} [{create_date}]({url})\ntweet_embed("{url}")\n\n')) %>%
  pull(formatted) %>% 
  purrr::walk(cat, sep="\n")
```
```{r kindle-highlights, eval=FALSE, include = FALSE}
kh <- readr::read_csv("~/projects/kindle-highlights/output.csv",
                      col_types = cols()) %>%
  mutate(formatted = paste0("\n\n> ", content, "  \n", "-- ", author, ", ", title)) %>%
  filter(between(as.Date(create_dt), end_date - 7, end_date)) 
kh %>%
  pull(formatted) %>%
  purrr::walk(cat, sep = "\n")
```

<!--
Harvest:
1.  Chrome history
2.  NYT
3.  Netflix, Amazon Prime
4.  Kindle highlights
-->







Week in review, `r start_date` - `r end_date`.

<!--more--> 
<script src="/rmarkdown-libs/twitter-widget/widgets.js"></script>



### Cooked

1.  Sweetcorn curry. ★★★  
    A save of the busted tomaticán recipe from the Chilean Kitchen.

### Seen

1.  Haikyu!!. ★★★  
    Need to find another good competition manga after this is done.

### Listened


### Read

1.  https://www.nytimes.com/interactive/2021/06/28/magazine/levar-burton-interview.html?action=click&module=Editors%20Picks&pgtype=Homepage
2.  https://www.nytimes.com/interactive/2021/06/27/us/louisiana-graves-enslaved-people.html?action=click&module=Top%20Stories&pgtype=Homepage 
3.  https://www.theinsight.org/p/rare-events-need-different-methods

### Upcoming

1.  https://bethduckles.podia.com/qualitative-research-for-quantitative-researchers

### To read
<!--starts at 30-->

8. "Thread! And for a vivid description of a wet bulb crisis in fiction, see Kim Stanley Robinson's Ministry For The Future.. https://t.co/9xHXUxZXNE," --timoreilly [Jun 29](https://twitter.com/2384071/status/1409689228234018816)

### To listen

1.  Ologies
6. "I just binged every available episode of Now &amp; Then by @HC_Richardson &amp; @jbf1755while doing touch-up trim painting. Fabulous history presented in these podcasts that ties the past to what's happening today. https://t.co/iFYRvBroPF," --maximum_mew [Jun 28](https://twitter.com/16494635/status/1409635858303447041)



### For study
3. "1) Wrote about shade inequality w/ @aliyaute and how poorer communities of color in cities like #Houston face hotter temperatures because of a lack of #trees 🌳. https://t.co/QvnS2uKrm0," --salamdogmillion [Jun 28](https://twitter.com/51544489/status/1409523115357523974)
7. "Communities that couldn't get mortgage loans in the 1940s are the same areas without fast home internet service today. There's no easy fix.. https://t.co/foa3ZcOIl4," --journojoshua [Jun 28](https://twitter.com/757955095086542848/status/1409647401376174080)
4. "thank you all for the thoughtful replies. some of the more interesting topics require a bit of background, so i have decided to start with something introductory, which we will build upon. https://t.co/bU8YZaaaA5"Variance Reduction in Monte Carlo Methods: Bilinear Strategies". https://t.co/278q97Vhed," --sam_power_825 [Jun 28](https://twitter.com/949284328315367425/status/1409561523832475650)


### For thought
10. "I think hedgefunds are the most suited to remote work because there's a lot of "daily action", which provides alternative fuel for bonding, normally powered by Duchenne smiles. And the pre-PMF startups are the worst hit, because the coworker Duchenne is all they have.," --danielgross [Jun 29](https://twitter.com/38190583/status/1409861159910621186)
5. "It is important to consider company stage when thinking about how to intervene to up your product game. What "works" in a big company will not work in a startup.At @Amplitude_HQ we tailor our coaching/feedback to meet companies where they are.(a quick morning activity ⬇️). https://t.co/UGMFiHRqEO," --johncutlefish [Jun 28](https://twitter.com/533409964/status/1409580615368773634)
1. "We can't keep regulating AI as if it..works. Most policy interventions start with the assumption that the technology lives up to its claims of performance but policymakers &amp; critical scholars need to stop falling for the corporate hype and should scrutinize these claims more.," --rajiinio [Jun 28](https://twitter.com/985158406125375488/status/1409333845795622912)






<!--### Project seeds, courses-->

<!--### To watch-->




### Pretties

9. "NEW POST: A collection of some of the best data visualisation work, articles, and other references discovered during April 2021. https://t.co/UxeRuqKSpZFeaturing work by @CraigTaylorViz @dr_xeo @framoriniii &amp; @_kathchan_. https://t.co/Kna2UCllJu," --visualisingdata [Jun 29](https://twitter.com/137712607/status/1409822362141343745)


```{r pretties, echo = FALSE}
# 2. "the department of safety and security at my girlfriend’s school just sent all an email telling students not to make eye contact with crows. https://t.co/GD4GbYV0LG," --gabrielledrolet [Jun 28](https://twitter.com/829155743215546372/status/1409508425738883073)
tweet_embed("https://twitter.com/829155743215546372/status/1409508425738883073")
```


<!--### Poem-->


